---
title: "QP1: King County Housing Price Analysis"
author: "Eric Hestekin"
date: "1/23/2020"
output: 
  html_document:
    number_sections: yes
    toc: yes
    toc_depth: 3
    toc_float: yes
---

Repo:  https://github.com/ehestekin/MSBA_5210_DataVis

``` {r echo = FALSE}
# Course: 5210 Communicating Data
# Purpose: Analyze king county housing prices
# Date: 1/23/20
# Author: Eric Hestekin
# Repo:  https://github.com/ehestekin/MSBA_5210_DataVis
```

``` {r echo = FALSE, include = FALSE}
# Clear environment of variables and functions
rm(list = ls(all = TRUE)) 

# Clear environmet of packages
if(is.null(sessionInfo()$otherPkgs) == FALSE)lapply(paste("package:", names(sessionInfo()$otherPkgs), sep=""), detach, character.only = TRUE, unload = TRUE)

```

``` {r setup, include = FALSE, warning = FALSE}
# Load Library
library(tidyverse)
library(gridExtra)
library(janitor)
library(knitr)

```

Load data and get initial summary
=================================
_Uni-variate non-graphical_

```{r}
#load data and print column names and first few rows
house_data <- read.csv('qp1_data.csv')

head(house_data, 5)
#get stats of basic housing quantitative data
summary(house_data[,1:5])
#categorical data
summary(house_data[,c('view','grade','condition','yr_built')])
#waterfront frequency
num_wf <- sum(house_data$waterfront)
num_wf
#percentage
round(num_wf / nrow(house_data) * 100, 2)

#sort by zip
house_data_by_zip <- house_data %>% group_by(zip_code) %>% 
  summarize_all(.funs = mean)

#add count.  couldn't figure out how to do this along with summarize_all
count <- tally(group_by(house_data, zip_code))

house_data_by_zip <- left_join(house_data_by_zip, count, by = "zip_code")

# Get 5 cheapest and most expensive zip codes
house_data_by_zip <- house_data_by_zip %>% arrange(price)

five_cheapest_zips = house_data_by_zip[1:5,'zip_code']
num_zips <- nrow(house_data_by_zip)
five_expense_zips  = house_data_by_zip[(num_zips-4):num_zips,'zip_code']

```

### Comments

+ Large range of house size and price (expected)
  + Right skewed
  
+ Not very many houses have a view.  Only upper quarter.
  + Similarly less than 1% are waterfront

+ Grade and condition both mostly symmetrically distributed
  + Median value for both is the middle rating.  No major bias.
  + Slight right skew.  Maybe due to average house being sold is above average   condition at time of sale?
  
+ Year is slightly left skewed.  Implies many older homes being sold.

### Questions

+ Location (zip code) effect on price for similar size house

+ Sqft vs number of bed/bath.  Size of yard effect?
  + I think size of yard would be sqft_lot - sqft_house / num_floors

+ Effect of year on price as well as renovation year

+ Shape of relationship between condition and price

Uni-variate graphical analysis
==============================

``` {r}

ggplot(house_data, mapping = aes(x = price)) +
  geom_histogram(binwidth = 50000)

#zoom on houses < 1.5mil
ggplot(house_data, mapping = aes(x = price)) +
  geom_histogram(binwidth = 50000) +
  xlim(0, 1.5e6) +
  ggtitle('King Co. Home Price Distribution', subtitle = 'Houses under 1.5mil only')

# house with view distro
ggplot(house_data, mapping = aes(x = view)) +
  geom_histogram(binwidth = 1)

#count houses with view (similar to waterfront above)
count(house_data, vars = view)

#year built distro
ggplot(house_data, mapping = aes(x = yr_built)) +
  geom_histogram(binwidth = 2) + 
  ggtitle('King Co. House Sale Home Build Year')

#house and yard sizes
# add yard size variable.  subtract footprint of house
# use floor function on floors (ha) to account for split levels
house_data <- house_data %>% mutate(sqft_yard = sqft_lot - (sqft_living / floor(floors)))

ggplot(house_data, mapping = aes(x = sqft_living)) +
  geom_histogram(binwidth = 50)

ggplot(house_data, mapping = aes(x = sqft_lot)) +
  geom_histogram(binwidth = 500) +
  xlim(0,5e4)

#look at price/sqft lot
ggplot(house_data, mapping = aes(x = price / sqft_lot)) +
  geom_histogram(binwidth = 10) +
  xlim(0,500)

```

### Comments

+ Price and Price per sqft have similar distributions

+ Only `r round(19489/21613, 2) *100`% of homes have a view

+ More newer houses than old which makes sense
  + Can see a major downturn around 1930 (great depression) and similar dips in 1970 and 2010 (recessions).  Makes sense fewer homes sold from those times as fewer homes were being built.
  
+ Square footage distributions look similar to price distros

### Questions

+ How closely related are price and square footage?

+ Impact of rare view or waterfront on price

+ Impact of age of home and how much remodel effects

Non Graphical Multivariate Analysis
===================================

``` {r}

# first test correlations of price with various factors

house_data %>% select(price,sqft_living, sqft_lot) %>% 
  cor() %>% round(3) %>% kable()

house_data %>% select(price, sqft_living, yr_built) %>% 
  cor() %>% round(3) %>% kable()

# condition vs age
house_data %>% select(yr_built, condition) %>% 
  cor() %>% round(3) %>% kable()

```

### Comments

+ Strong correlation between price and living sq ft
  + Not much of a correlation for lot size
  
+ Age and price not directly correlated.
  + Age and square footage have a stronger correlation.
    + Implies houses have been getting bigger as time goes on
    
+ Negative correlation between condition and year which makes sense
  + Same as a positve correlation between _age_ and condition
  
Multivariate Graphical Analysis
===============================

``` {r}

# continue looking into correlated variables
ggplot(house_data, mapping = aes(sqft_living, price)) +
  geom_point()

#effect of view
view_group <- house_data %>% group_by(view) %>% summarize(avg = mean(price))

ggplot(view_group, mapping = aes(x = view, y = avg)) +
  geom_bar(stat = 'identity')

#cool alternate way without doing group_by into a new df
ggplot(house_data, mapping = aes(x = view, y  = price)) +
  geom_bar(stat = 'summary')

#plot price per sqft and add color for most expensive and least expensive zips
#first make subset of data
house_data_5_expense_zips <- house_data[house_data$zip_code %in% five_expense_zips$zip_code,]
ggplot(house_data,mapping = aes(sqft_living, price)) +
  geom_point() +
  geom_point(house_data_5_expense_zips, mapping = aes(sqft_living, price),color = 'red')

#try similar idea with histogram
ggplot(house_data, mapping = aes(x = price)) +
  geom_histogram(binwidth = 50000) +
  xlim(0, 1.5e6) +
  geom_histogram(data = house_data_5_expense_zips, mapping = aes(price)
                 , fill = 'red', binwidth = 50000)

#would need to normalize above plot because it is hard to see.  good idea for later

#condition vs price
ggplot(house_data, mapping = aes(x = condition, y = price / sqft_living)) +
  geom_point() +
  geom_point(house_data_5_expense_zips, mapping = aes(x = condition, y = price / sqft_living), color = 'red')

ggplot(house_data, mapping = aes(sqft_living, price)) +
  geom_point() +
  geom_smooth(method = 'lm')#do smooth for each condition, see if slopes are same

```

Refine Plots
============

```` {R}

# view vs price

ggplot(house_data, mapping = aes(x = view, y  = price)) +
  geom_bar(stat = 'summary', color = 'black', fill = c('#996666','#00aa78','#00aa78','#00aa78','#00aa78')) +
  xlab('View Rating') +
  ggtitle('Effect of View on House Prices') +
  ylab('Average Price ($)')

ggsave(filename = 'view_vs_price.png', plot = last_plot())

# effect of age and renovations
p_age <- ggplot(house_data, mapping = aes(x = yr_built, y = price / sqft_living)) +
          geom_point(alpha = 0.1) + 
          ggtitle('King Co. Home Build Year vs Price per SqFt') +
          xlab('Home Build Year') +
          ylab('Price per SqFt ($)')

#add regression for year and rennovation
build_yr_lm <- lm(price / sqft_living ~ yr_built, data = house_data)
renov_only_data <- house_data %>% filter(yr_renovated > 0)
renov_lm <- lm(price / sqft_living ~ yr_renovated, data = renov_only_data)

#p_age <- p_age + geom_abline(slope = build_yr_lm$coefficients[2], intercept = build_yr_lm$coefficients[1], col = '#0000cc')

#look at rennovation year with condition
p_age <-  p_age + geom_abline(slope = renov_lm$coefficients[2], intercept = renov_lm$coefficients[1], col = '#cc0000')

ggsave(filename = 'price_vs_yr_built.png', plot = p_age)

# condition
ggplot(house_data, mapping = aes(x= condition, y = price)) +
  geom_bar(stat = 'summary', color = 'black', fill = c('#996666','#996666','#00aa78','#00aa78','#00aa78')) +
  xlab('Condition Rating') +
  ylab('Average Sale Price ($)') +
  ggtitle('House Condition vs Sale Price')
  
  ggsave(filename = 'condition_vs_price.png', plot = last_plot())

```