---
title: 'PA1 OJ: Minute Maid vs Citrus Hill Purchases'
output:
  html_document:
    number_sections: yes
    toc: yes
    toc_depth: 3
    toc_float: yes
  word_document:
    toc: yes
    toc_depth: '3'
date: "1/9/20"
---

Repo:  https://github.com/ehestekin/MSBA_5210_DataVis

``` {r echo = FALSE}
# Course: 5210 Communicating Data
# Purpose: Analyze orange juice purchases
# Date: 1/9/20
# Author: Eric Hestekin
# Repo:  https://github.com/ehestekin/MSBA_5210_DataVis
```

``` {r echo = FALSE, include = FALSE}
# Clear environment of variables and functions
rm(list = ls(all = TRUE)) 

# Clear environmet of packages
if(is.null(sessionInfo()$otherPkgs) == FALSE)lapply(paste("package:", names(sessionInfo()$otherPkgs), sep=""), detach, character.only = TRUE, unload = TRUE)


```

``` {r setup, include = FALSE, warning = FALSE}
# Load Library
library(tidyverse)
library(gridExtra)

```

``` {r}
# Load Data
oj_data <- read.csv("OJ_Data.csv")

# View first few rows of data
head(oj_data, 3)
```

Are the distributions symmetric or skewed?
==========================================

``` {r} 

summary(oj_data)

```

+ The distributions of price are very symmetric.
+ The distributions of discounts are not symmetric.
  + This is likely explained by how infrequently discounts are offered.
+ Citrus Hill accounted for `r round(summary(oj_data$Purchase[oj_data$Purchase == 'CH'])[1]/dim(oj_data)[1] * 100 , 2)  ` % of sales

+ Skewness affects whether we can use a normal regression (need normal dist) or not (when it is skewed)

Which brand sells the most?
===========================

+ Citrus Hill sells more (61% of sales)
+ This may be due to price or ammount of discounts

``` {r}

# add column for which price is lower (must account for discount)
oj_data <- mutate(oj_data, ch_lower_price = (PriceCH - DiscCH) < (PriceMM - DiscMM))

#test affect of lower price on purchase TODO

```

Which store sells the most OJ?
==============================

``` {r}

#count occurences of each store
per_store_sales <- count(oj_data, STORE)

```
+ Store `r per_store_sales$STORE[per_store_sales$n == max(per_store_sales$n)]` sold the most OJ with `r max(per_store_sales$n)` sales

More variables to add?
======================

+ Above in the 'Which brand sells the most' section, I added a column for which brand had a lower price for a given sale datapoint.
+ It would also be good to filter down discounts to only where discounts applied to get a better idea of the effect they have

Other comments/questions from EDA of data?
==========================================

+ It would be interesting to know the percent of sales of each brand at each store.
+ It would also be interesting to see if the price/discounts has a different effect at different stores.  Maybe some stores have less price-sensative shoppers.

Graphical Analysis
==================

``` {R, warning = FALSE}

grid.arrange(
ggplot(data = oj_data, mapping = aes(x = PriceMM - DiscMM)) +
  geom_histogram(binwidth = 0.02, fill = 'orange') +
  xlim(1.65,2.3) +
  ylim(-25, 300) + 
  xlab('Minute Maid Sales Price ($)'),
# binwidth of 2cents for prices

ggplot(data = oj_data, mapping = aes(x = PriceCH - DiscCH)) + 
  geom_histogram(binwidth = 0.02, fill = 'orange') +
  xlim(1.65,2.3) +
  ylim(-25, 300) +
  xlab('Citrus Hill Sales Price ($)'),
ncol = 1)

```

+ Distribution of prices for Citrus Hill significantly lower than Minute Maid

``` {R}

#add similar box plot to clarify distributions
grid.arrange(
ggplot(data = oj_data, mapping = aes(x = 1)) +
  geom_boxplot(mapping = aes(y = PriceMM - DiscMM), fill = 'orange') +
  coord_flip() + 
  ylim(1.65,2.3) +
  xlab(''),

ggplot(data = oj_data, mapping = aes(x = 1)) +
  geom_boxplot(mapping = aes(y = PriceCH - DiscCH), fill = 'orange') + 
  coord_flip() + 
  ylim(1.65,2.3) +
  xlab(''),

ncol = 1)
```

+ Confirm average price and overall distribution of price is lower for Citris Hill