---
title: "USA Domestic Flights Growth Analysis"
author: "Eric Hestekin"
date: "3/19/2020"
output: slidy_presentation
---


``` {r echo = FALSE}
# Course: 5210 Communicating Data
# Purpose: USA Domestic Flights Growth Analysis
# Date: 3/19/20
# Author: Eric Hestekin
# Repo:  https://github.com/ehestekin/MSBA_5210_DataVis
```

``` {r echo = FALSE, include = FALSE}
# Clear environment of variables and functions
rm(list = ls(all = TRUE)) 
# Clear environmet of packages
if(is.null(sessionInfo()$otherPkgs) == FALSE)lapply(paste("package:", names(sessionInfo()$otherPkgs), sep=""), detach, character.only = TRUE, unload = TRUE)
```

``` {r setup, include = FALSE, warning = FALSE}
# Load Library
library(tidyverse)
library(scales)
library(gridExtra)
library(inspectdf)
library(ggrepel)
```


```{r, include = FALSE, warning = FALSE}
#Load Data 
yrly_com_flights_by_city <- read_csv('yrly_com_flights_by_city.csv')
com_flights_df <- read_csv('com_flights_df.csv')
```


Recent air travel market downturn
====================================

```{r, echo = FALSE}
pop_trend_data <- com_flights_df %>% 
  group_by(year) %>% 
  summarise(tot_passenger_vol = sum(passengers))

pop_trend_plot <- ggplot(pop_trend_data, mapping = aes(x = year, y = tot_passenger_vol/1e6)) +
  geom_line(color = 'grey') +
  geom_line(filter(pop_trend_data, year > 2006), 
            mapping = aes(x = year, y = tot_passenger_vol/1e6),
            size = 1.5, 
            color = '#750000')

pop_trend_plot <- pop_trend_plot + 
  geom_point(data = filter(pop_trend_data, year == 2009), 
             mapping = aes(x = year, y = tot_passenger_vol/1e6), 
             fill = '#344182', shape = 25, size = 3) +
  geom_text(aes(x = 2010, y = tot_passenger_vol[year == 2009]/1e6 - 10), label = '526.3M', color = '#344182' )


pop_trend_plot + theme_classic() +
  theme(axis.line.y = element_blank(),
        axis.line.x = element_blank(),
        axis.title.x  = element_blank(),
        plot.title = element_text(color = "#808080",
                                  size = 16, face = "bold"),
        plot.subtitle = element_text(color = "#808080", 
                                     size = 12, face = "plain"),
        axis.text.x = element_text(color = "#808080", size = 11),
        axis.text.y = element_text(color = "#808080", size = 11),
        axis.title = element_text(color = "#808080", size = 11, face = "plain"),
        legend.title = element_text(color = "#808080", size = 12, face = "plain"),
        legend.text = element_text(color = "#808080", size = 12, face = "plain"),
        plot.caption = element_text(color = "#808080", size = 10, face = "plain", margin = margin(t=10))
        ) +
  scale_x_continuous(breaks = seq(2005,2010, by = 2)) +
  labs(title = 'Total passenger volume is falling',
       subtitle = expression(paste('Down ',
                                   italic(bold('9.1%')),
                                   ' from 2007 peak')),
       y = 'Millions of Passengers')
```

 - Sharp decrease in passenger volume presents revenue challenges

<font size = '3'>Source Data:  [US domestic flights from 1990 to 2009, US Census Bureau](http://academictorrents.com/details/a2ccf94bbb4af222bf8e69dad60a68a29f310d9a)</font>

Despite volume drop, operating efficiency is up 
===============================================

```{r, echo = FALSE}
filled_trend_data <- com_flights_df %>% 
  group_by(year) %>% 
  summarise(avg_load_factor = mean(load_factor))

filled_trend_plot <- ggplot(filled_trend_data, 
                            mapping = aes(x = year,
                                          y = avg_load_factor/100)) +
  geom_line(color = 'grey') +
  geom_line(filter(filled_trend_data, year > 2006), 
            mapping = aes(x = year, y = avg_load_factor/100),
            size = 1.5, 
            color = '#344182')

filled_trend_plot <- filled_trend_plot + 
  geom_point(data = filter(filled_trend_data, year == 2009), 
             mapping = aes(x = year, y = avg_load_factor/100), 
             fill = '#344182', shape = 25, size = 3) +
  geom_text(aes(x = 2010, y = avg_load_factor[year == 2009]/100 - 0.01), label = '71%', color = '#344182' )


filled_trend_plot + theme_classic() +
  theme(axis.line.y = element_blank(),
        axis.line.x = element_blank(),
        axis.title  = element_blank(),
        plot.title = element_text(color = "#808080",
                                  size = 16, face = "bold"),
        plot.subtitle = element_text(color = "#808080", 
                                     size = 12, face = "plain"),
        axis.text.x = element_text(color = "#808080", size = 11),
        axis.text.y = element_text(color = "#808080", size = 11),
        legend.title = element_text(color = "#808080", size = 12, face = "plain"),
        legend.text = element_text(color = "#808080", size = 12, face = "plain"),
        plot.caption = element_text(color = "#808080", size = 10, face = "plain", margin = margin(t=10))
        ) +
  scale_x_continuous(breaks = seq(2005,2010, by = 2)) +
  scale_y_continuous(labels = percent_format()) +
  labs(title = 'However, percent of seats filled has gone up',
       subtitle = 'Up 1 percentage point from 2007 total passenger peak')
```

- With fewer passengers it will be important to manage number of seats offered
- Empty seats eat up profit

Growing cities are a good bet for successful expansion
======================================================

```{r, echo = FALSE}
yrly_com_flights_by_city %>% 
  ggplot(mapping = aes(x = total_pop_growth/1000000, y = total_passengers/1000000)) +
  geom_point(color = 'grey') + 
  geom_smooth(color = '#344182', method = 'lm', se = FALSE) +
  theme_classic() +
  theme(axis.line.y = element_blank(),
        axis.line.x = element_blank(),
        axis.title  = element_text(color = "#808080", size = 11, face = "bold"),
        plot.title = element_text(color = "#808080",
                                  size = 16, face = "bold"),
        plot.subtitle = element_text(color = "#808080", 
                                     size = 12, face = "plain"),
        axis.text.x = element_text(color = "#808080", size = 11),
        axis.text.y = element_text(color = "#808080", size = 11),
        legend.title = element_text(color = "#808080", size = 12, face = "plain"),
        legend.text = element_text(color = "#808080", size = 12, face = "plain"),
        plot.caption = element_text(color = "#808080", size = 10, face = "plain", margin = margin(t=10))
        ) +
  xlab('Population growth') +
  ylab('Total passengers') +
  labs(title = 'Population growth correlates strongly with passenger growth',
       subtitle = 'Numbers for previous five years (in millions)\nDallas, TX fastest growing city (by far)') +
  geom_label_repel(filter(yrly_com_flights_by_city, origin_city == 'Dallas, TX'), 
                   mapping = aes(x = total_pop_growth/1000000, y = total_passengers/1000000, label = origin_city),
                   segment.color = '#344182', min.segment.length = 0, nudge_y = -15.5, label.size = NA, color = '#344182')
```

+ Despite massive population growth in previous five years, Dallas did not have growth in passenger volume

High load factor locations provide growth opportunity too
=====================================================================================
```{r, echo = FALSE}
yrly_com_flights_by_city %>% 
  filter(percent_pop_growth > 0) %>% 
  filter(total_load_factor > 0.65) %>% 
  ggplot(mapping = aes(x = total_load_factor, y = percent_pop_growth/100)) +
  geom_point(color = 'grey') +
  theme_classic() +
  theme(axis.line.y = element_blank(),
        axis.line.x = element_blank(),
        axis.title  = element_text(color = "#808080", size = 11, face = "bold"),
        plot.title = element_text(color = "#808080",
                                  size = 16, face = "bold"),
        plot.subtitle = element_text(color = "#808080", 
                                     size = 12, face = "plain"),
        axis.text.x = element_text(color = "#808080", size = 11),
        axis.text.y = element_text(color = "#808080", size = 11),
        legend.title = element_text(color = "#808080", size = 12, face = "plain"),
        legend.text = element_text(color = "#808080", size = 12, face = "plain"),
        plot.caption = element_text(color = "#808080", size = 10, face = "plain", margin = margin(t=10))
        ) +
  scale_x_continuous(labels = percent_format()) +
  scale_y_continuous(labels = percent_format()) +
  xlab('Load factor') +
  ylab('Population growth') +
  labs(title = 'Seattle and Miami are also prime candidates for expansion',
       subtitle = 'High load factor, above average population growth') +
  geom_label_repel(filter(yrly_com_flights_by_city, origin_city %in% c('Charlotte, NC', 'Seattle, WA', 'Dallas, TX', 'Miami, FL')), 
                   mapping = aes(x = total_load_factor, y = percent_pop_growth/100, label = origin_city),
                   segment.color = '#344182', min.segment.length = 0, nudge_y = 0.05, nudge_x = 0.05, label.size = NA, color = '#344182')
```

+ Above average load factor implies more seats could be offered profitably.
+ Combined with positive population growth makes these cities the best bet

Charlotte should be priority #1 for further growth
==================================================
```{r, echo = FALSE}
# all cities with positve passenger growth and positive load factor growth
yrly_com_flights_by_city %>% 
  filter(total_passenger_growth > 0, avg_load_factor_growth > 0) %>% 
  arrange(desc(total_passenger_growth)) %>% 
  ggplot(mapping = aes(x = avg_load_factor_growth/100, y = total_passenger_growth/1e6, label = origin_city)) + 
  geom_point(color = 'grey') + 
  theme_classic() +
  theme(axis.line.y = element_blank(),
        axis.line.x = element_blank(),
        axis.title  = element_text(color = "#808080", size = 11, face = "bold"),
        plot.title = element_text(color = "#808080",
                                  size = 16, face = "bold"),
        plot.subtitle = element_text(color = "#808080", 
                                     size = 12, face = "plain"),
        axis.text.x = element_text(color = "#808080", size = 11),
        axis.text.y = element_text(color = "#808080", size = 11),
        legend.title = element_text(color = "#808080", size = 12, face = "plain"),
        legend.text = element_text(color = "#808080", size = 12, face = "plain"),
        plot.caption = element_text(color = "#808080", size = 10, face = "plain", margin = margin(t=10))
        ) +
  scale_x_continuous(labels = percent_format()) +
  xlab('Annualized load factor growth') +
  ylab('Passenger growth (in millions)') +
  labs(title = 'Charlotte is a major positive growth outlier',
       subtitle = 'Both load factor growth and passenger growth are strong\n(only positve passenger growth cities shown) ') +
  geom_label_repel(filter(yrly_com_flights_by_city, origin_city %in% c('Charlotte, NC', 'Seattle, WA', 'Miami, FL')), 
                   mapping = aes(x = avg_load_factor_growth/100, y = total_passenger_growth/1e6, label = origin_city),
                   segment.color = '#344182', min.segment.length = 0, nudge_y = -.15, label.size = NA, color = '#344182')

```

Conclusion
==============

- Decreasing passenger volumes make efficiency improvements critical
- Dallas, TX is the country's fastest growing city by population, but has not yet had more seats offered to take advantage of larger population
- Seattle and Miami are good candidates for growth due to:
    - Growing population
    - Increasing passenger numbers despite wider industry downturn
    - High load factor lowers risk
- Charlotte has the most impressive recent growth and should remain a priority for further growth

<font size = '3'>
For comments or questions:  
Eric Hestekin  
ehestekin@seattleu.edu
</font>
