---
title: "Airport Growth Potential"
author: "Eric Hestekin"
date: "2/27/2020"
output: 
  html_document:
    number_sections: yes
    toc: yes
    toc_depth: 3
    toc_float: yes
---

Repo:  https://github.com/ehestekin/MSBA_5210_DataVis

``` {r echo = FALSE}
# Course: 5210 Communicating Data
# Purpose: Final Project: Airport Growth Analysis
# Date: 2/13/20
# Author: Eric Hestekin, Schyuler Lujan
# Repo:  https://github.com/ehestekin/MSBA_5210_DataVis
```

``` {r echo = FALSE, include = FALSE}
# Clear environment of variables and functions
rm(list = ls(all = TRUE)) 
# Clear environmet of packages
if(is.null(sessionInfo()$otherPkgs) == FALSE)lapply(paste("package:", names(sessionInfo()$otherPkgs), sep=""), detach, character.only = TRUE, unload = TRUE)
```

``` {r setup, include = FALSE, warning = FALSE}
# Load Library
library(tidyverse)
library(scales)
library(gridExtra)
library(inspectdf)
library(ggrepel)
```


# Data Load and Wrangle

Data source :  [US domestic flights from 1990 to 2009, US Census Bureau](http://academictorrents.com/details/a2ccf94bbb4af222bf8e69dad60a68a29f310d9a)

```{r, message = FALSE, warning = FALSE}
#Load Data TSV
#no column headers so set array of names
col_labels = c('origin_code',
               'dest_code',
               'origin_city',
               'dest_city',
               'passengers',
               'seats',
               'num_flights',
               'distance',
               'year_month',
               'origin_pop',
               'dest_pop')

#suppress colum spec message.  (Markdown code block setting doesn't work)
flights_df <- suppressMessages(read_tsv('flight_edges.tsv',col_names = col_labels))
```

Goal is to identify airports with growth potential
+ Answer the business question "Which airports should be targets for expansion?"
+ What airport and city characteristics are good predictors of flight growth?

## First look at data

```{r}
head(flights_df)

summary(flights_df)
```

## Wrangle

```{r}
#separate month_year
#first 4 chars are year, convert from string to num
flights_df <- flights_df %>% 
  separate(year_month, into = c('year','month'), sep = 4, convert = TRUE) 

#add percentage of seats filled for each flight
#and remove rows where seats are zero (no flight data)
# and also remove flights where seat/flight < 4 (non commercial flight)
com_flights_df <- flights_df %>%
  mutate(load_factor = (passengers/seats) * 100) %>% 
  filter(seats != 0) %>%
  filter(passengers != 0) %>% 
  filter(seats/num_flights > 4) %>% 
  #focus on routes offering >500 seats/wk (2150/month)
  #way too many routes with a very small number of people flying
  filter(seats > 2150)

#also throw out seat filled percentages > 100
#this shouldn't be possible so don't trust data point
com_flights_df <- filter(com_flights_df, load_factor <= 100)

#focusing on current growth prospects so only keep previous 5 years of data
com_flights_df <- com_flights_df %>% 
  filter(year > (max(year) - 5))

#calculate population increase year to year
pop_increases <- com_flights_df %>%
  group_by(origin_city, year) %>% 
  summarize(avg_pop = mean(origin_pop)) %>% 
  arrange(desc(year), .by_group = TRUE) %>% 
  mutate(pop_increase = avg_pop - lead(avg_pop))

#calculate route seats flown and offered increase year to year
flight_increases <- com_flights_df %>% 
  group_by(origin_code, dest_code, year) %>% 
  summarize(yrly_passenger = sum(passengers), yrly_seats = sum(seats)) %>% 
  arrange(desc(year), .by_group = TRUE) %>% 
  mutate(pass_growth = yrly_passenger - lead(yrly_passenger), seat_growth = yrly_seats - lead(yrly_seats), 
         yrly_load_factor = yrly_passenger / yrly_seats * 100, load_factor_growth = yrly_load_factor - lead(yrly_load_factor))

#put yrly data together
#first add full city names
#then add population numbers
yrly_com_flights_df <- inner_join(flight_increases, distinct(select(com_flights_df, origin_code, origin_city)), by = 'origin_code') %>% 
  inner_join(distinct(select(com_flights_df, dest_code, dest_city)), by = 'dest_code') %>% 
  inner_join(pop_increases, by = c('year', 'origin_city')) %>%
  #rename columns for origin city so able to add destination population data
  rename(origin_pop = avg_pop, origin_pop_increase = pop_increase) %>% 
  inner_join(rename(pop_increases, dest_city = origin_city, dest_pop = avg_pop, dest_pop_increase = pop_increase), by = c('year', 'dest_city')) %>% 
  #distance got dropped during grouping
  inner_join(distinct(select(com_flights_df, origin_code, dest_code, distance)), by = c('origin_code', 'dest_code'))

#noticed there are duplicates of some rows (everything is same except distance is one mile different)
yrly_com_flights_df_no_dup_test <-  yrly_com_flights_df %>% distinct(origin_code, dest_code, year, .keep_all = TRUE)

#noticed the duplicates with the CLT/ABE route in the Individual Route and City Analysis section
#check dupes were removed
yrly_com_flights_df_no_dup_test %>% filter(origin_code == 'CLT')

#looks like it worked so save df
yrly_com_flights_df <- yrly_com_flights_df_no_dup_test
rm(yrly_com_flights_df_no_dup_test)
  
# Save df for use in PowerBI Dash
write_csv(yrly_com_flights_df, path = 'yrly_com_flights_df.csv')


#get summary now with tidy commercial flights data set
summary(yrly_com_flights_df)
```

*Notes*

+ Would be useful to look at averages for passengers and seats on a single route
+ Interesting distribution of distances
    + Looks like many short routes (median << mean)
+ Population skewed towards very large cities (due to more flight routes?)
+ Fairly symmetrical load factor
+ NAs result from not having previous data
    + New route or first year in data (2005)
    
+ Populations are growing but distribution is skewed (mean>>median)

**Assuming very minimal net one-way travelers so origin/destination numbers are symmetric**
+ Focus on departing flights (origin city sorting) and assume numbers are nearly identical for arrivals

```{r}
# Quick check above assumption with Chicago (busiest airport by passenger volume, found below)
yrly_com_flights_df %>% filter(origin_code == 'ORD') %>% 
  group_by(year) %>% 
  summarize(total_departs = sum(yrly_passenger)) / #divide the arrives by departs

yrly_com_flights_df %>% filter(dest_code == 'ORD') %>% 
  group_by(year) %>% 
  summarize(total_arrives = sum(yrly_passenger))
```
+ All years at ORD are within 0.4% of equal departs and arrivals

# Graphical Analysis

## How has data changed over time covered in dataset?
```{r}
com_flights_df %>% 
  group_by(year) %>% 
  summarise(avg_num_flights = mean(num_flights)) %>% 
  ggplot(mapping = aes(x = year, y = avg_num_flights)) +
  geom_line() +
  geom_point()

com_flights_df %>% 
  group_by(year) %>% 
  summarise(avg_load_factor = mean(load_factor)) %>% 
  ggplot(mapping = aes(x = year, y = avg_load_factor)) +
  geom_line() +
  geom_point()

com_flights_df %>% 
  group_by(year) %>% 
  summarise(avg_passenger_vol = mean(passengers)) %>% 
  ggplot(mapping = aes(x = year, y = avg_passenger_vol)) +
  geom_line() +
  geom_point()

com_flights_df %>% 
  group_by(year) %>% 
  summarise(tot_passenger_vol = sum(passengers)) %>% 
  ggplot(mapping = aes(x = year, y = tot_passenger_vol)) +
  geom_line() +
  geom_point()
```

+ Drop in passenger volumes from 2007 - 2009
+ Increase in percentage of seats filled in same time period

## Look at individual variables
### Distance

```{r}
grid.arrange(
  ggplot(data = yrly_com_flights_df, mapping = aes(x = 1, y = distance)) +
    geom_boxplot() +
    coord_flip(),
  #100 mile bins for easier increments
  ggplot(data = yrly_com_flights_df, mapping = aes(x = distance)) +
    geom_histogram(binwidth = 100),
  
  ncol = 1
)
```

+ Domestic flights mostly are between 300 - 1000 miles

### Load Factor

```{r}
grid.arrange(
  ggplot(data = yrly_com_flights_df, mapping = aes(x = 1, y = yrly_load_factor)) +
    geom_boxplot() +
    coord_flip(),
  
  ggplot(data = yrly_com_flights_df, mapping = aes(x = yrly_load_factor)) +
    geom_histogram(bins = 50),
  
  ncol = 1
)
```

+ Vast majority of flights are above half full
+ Half of the routes were greater than ~75% full

### Passengers

```{r}
grid.arrange(
  ggplot(data = yrly_com_flights_df, mapping = aes(x = 1, y = yrly_passenger)) +
    geom_boxplot() +
    coord_flip(),
  
  ggplot(data = yrly_com_flights_df, mapping = aes(x = yrly_passenger)) +
    geom_histogram(bins = 50),
  
  ncol = 1
)
```

+ There are lots of routes with very few yearly passengers.
    + Median >> Mean
    
### Load Factor and Passenger Growth (last year)

```{r}
grid.arrange(
  ggplot(data = filter(yrly_com_flights_df, year == 2009), mapping = aes(x = 1, y = load_factor_growth)) +
    geom_boxplot() +
    coord_flip(),
  
  ggplot(data = filter(yrly_com_flights_df, year == 2009), mapping = aes(x = load_factor_growth)) +
    geom_histogram(bins = 50),
  
  ncol = 1
)
```

```{r}
grid.arrange(
  ggplot(data = filter(yrly_com_flights_df, year == 2009), mapping = aes(x = 1, y = pass_growth)) +
    geom_boxplot() +
    coord_flip(),
  
  ggplot(data = filter(yrly_com_flights_df, year == 2009), mapping = aes(x = pass_growth)) +
    geom_histogram(bins = 100),
  
  ncol = 1
)
```

+ Growth relatively symmetrical distribution
+ Median above zero, showing overall load factor growth
+ Median below zero, showing overall negative passenger growth
+ Interested to look at the large growth routes

# Individual Route and City Analysis

## Top Airports

+ Looking at cities instead of departure airports to make analysis consistent for places with multiple airports

### By Passenger Volume

```{r}

yrly_com_flights_df %>% group_by(origin_city) %>% 
  summarize(total_passengers = sum(yrly_passenger)) %>% 
  arrange(desc(total_passengers)) %>% 
  head(10)
```

### By Load Factor

```{r}

yrly_com_flights_df %>% group_by(origin_city) %>% 
  summarize(total_load_factor = sum(yrly_passenger)/ sum(yrly_seats)) %>% 
  arrange(desc(total_load_factor)) %>% 
  head(10)
```

### Five year passenger volume growth

```{r}

yrly_com_flights_df %>% group_by(origin_city) %>% 
  summarize(total_passenger_growth = sum(pass_growth, na.rm = TRUE)) %>% 
  arrange(desc(total_passenger_growth)) %>% 
  head(10)
```

### Five year seat volume growth

```{r}

yrly_com_flights_df %>% group_by(origin_city) %>% 
  summarize(total_seat_growth = sum(seat_growth, na.rm = TRUE)) %>% 
  arrange(desc(total_seat_growth)) %>% 
  head(10)
```

+ Interesting Charlotte has largest passenger growth but is 3rd in seat growth

+ *Does that imply a large load factor increase?*

```{r}
#take closer look at Charlotte
#EDIT:  Used this to identify duplicate ABE routes which greatly affected Charlotte numbers.  Fixed in wrangle section above
#yrly_com_flights_df %>% filter(origin_code == 'CLT')

```

### Five year load factor growth

```{r}
#total five year load factor growth
yrly_com_flights_df %>% group_by(origin_city) %>% 
  summarize(total_load_factor_growth = sum(load_factor_growth, na.rm = TRUE)) %>% 
  arrange(desc(total_load_factor_growth)) %>% 
  head(10)

#average growth per route
yrly_com_flights_df %>% group_by(origin_city) %>% 
  summarize(n = n(), avg_load_factor_growth = sum(load_factor_growth, na.rm = TRUE) / n) %>% 
  arrange(desc(avg_load_factor_growth)) %>% 
  head(10)

#above top 10 is all very small routes (n < 33 over five years)
#check with n > 100 (20 routes/yr)
yrly_com_flights_df %>% group_by(origin_city) %>% 
  summarize(n = n(), avg_load_factor_growth = sum(load_factor_growth, na.rm = TRUE) / n) %>%
  filter(n > 100) %>% 
  arrange(desc(avg_load_factor_growth)) %>% 
  head(10)


```

+ **As expected from passenger/seat numbers, Charlotte has highest load factor growth**

## City Growth Analysis

```{r}
#total five year pop growth
pop_increases %>% group_by(origin_city) %>% 
  summarize(total_pop_growth = sum(pop_increase, na.rm = TRUE)) %>% 
  arrange(desc(total_pop_growth)) 

#by percentage
pop_increases %>% group_by(origin_city) %>% 
  summarize(min_year = min(year), percent_pop_growth = sum(pop_increase, na.rm = TRUE) / avg_pop[year == min_year] * 100) %>% 
  arrange(desc(percent_pop_growth)) 
```

+ Austin and Charlotte are in top ten for load factor growth and percent population growth
+ Total population growth and passenger growth also have cities in common
+ Dallas has by far the largest population growth
    
+ **Correlation between city growth and airport growth?**


# Statistics

## Check correlations

```{r}
#this plot doesn't look right.  maybe not the right format of df
# yrly_com_flights_df %>% 
#   group_by(origin_city, year) %>% 
#   select(yrly_passenger, yrly_seats, origin_pop, dest_pop) %>% 
#   inspect_cor() %>% 
#   show_plot()

#try on one city
yrly_com_flights_df %>%
  filter(origin_city == 'Charlotte, NC') %>%
  select(yrly_passenger, yrly_seats, origin_pop, dest_pop) %>%
  inspect_cor() %>%
  show_plot()

```

+ Still not a fan of above method from week 10.  Do manual testing instead

### Population and Passengers

```{r}
#First group flights by city so individual routes aren't being tested
#combine all the above parameters into this dataset
yrly_com_flights_by_city <- yrly_com_flights_df %>% 
  group_by(origin_city) %>% 
  summarize(total_passengers = sum(yrly_passenger),
            total_seats = sum(yrly_seats),
            total_load_factor = sum(yrly_passenger)/ sum(yrly_seats),
            total_passenger_growth = sum(pass_growth, na.rm = TRUE),
            passenger_growth_percentage = total_passenger_growth / total_passengers * 100,
            total_seat_growth = sum(seat_growth, na.rm = TRUE),
            total_load_factor_growth = sum(load_factor_growth, na.rm = TRUE),
            n = n(), 
            avg_load_factor_growth = sum(load_factor_growth, na.rm = TRUE) / n)

#add pop data and join
pop_increases_by_city <- pop_increases %>% 
  group_by(origin_city) %>% 
  summarize(total_pop_growth = sum(pop_increase, na.rm = TRUE),
            min_year = min(year), 
            percent_pop_growth = sum(pop_increase, na.rm = TRUE) / avg_pop[year == min_year] * 100,
            avg_pop = mean(avg_pop))#keep population in df. 5yr avg in this case

#join
yrly_com_flights_by_city <- inner_join(yrly_com_flights_by_city, pop_increases_by_city, by = 'origin_city') %>% 
  select(-min_year) %>% 
  filter(n > 50) # also filter out very small cities (less than 10 routes/yr)

#save for potential use on dash
write_csv(yrly_com_flights_by_city, path = 'yrly_com_flights_by_city.csv')
            
```


```{r}
cor.test(yrly_com_flights_by_city$total_passengers, yrly_com_flights_by_city$avg_pop,
         method = 'pearson')
```

```{r, warning = FALSE}
#test inspect_cor again with grouped df
yrly_com_flights_by_city %>% 
  inspect_cor() %>% 
  show_plot()

#narrow down for easier reading
yrly_com_flights_by_city %>% 
  inspect_cor(with_col = c('total_pop_growth', 'percent_pop_growth')) %>% 
  show_plot()

```

+ High correlation between total population growth and total passengers
+ Not quite as high but still significant correlation between population growth and load factor growth
+ Interesting negative correlation between population growth and seat growth

**Want to identify cities with:**
- Population growth (correlated to passenger and load factor growth)
- High load factors (room for more seats as higher percentage is filled)


## Test significance of top cities from rankings above

```{r}
z <- qnorm(0.975)

yrly_com_flights_df %>% group_by(origin_city) %>% 
  summarize(n = n(), 
            avg_load_factor_growth = sum(load_factor_growth, na.rm = TRUE) / n,
            sd = sd(load_factor_growth , na.rm = TRUE) / n,
            ci = z*sd / sqrt(n)) %>%
  filter(n > 100) %>% 
  arrange(desc(avg_load_factor_growth)) %>% 
  head(10) %>% 
  ggplot(mapping = aes(x = reorder(origin_city, avg_load_factor_growth), y = avg_load_factor_growth)) +
  geom_bar(stat = 'identity') +
  geom_errorbar(aes(ymin = avg_load_factor_growth - ci, ymax = avg_load_factor_growth + ci)) +
  coord_flip()
```


```{r}
z <- qnorm(0.975)

yrly_com_flights_df %>% group_by(origin_city) %>% 
  summarize(n = n(), 
            total_load_factor_growth = sum(load_factor_growth, na.rm = TRUE) ,
            sd = sd(load_factor_growth, na.rm = TRUE),
            ci = z*sd / sqrt(n)) %>%
  filter(n > 100) %>% 
  arrange(desc(total_load_factor_growth)) %>% 
  head(10) %>% 
  ggplot(mapping = aes(x = reorder(origin_city, total_load_factor_growth), y = total_load_factor_growth)) +
  geom_bar(stat = 'identity') +
  geom_errorbar(aes(ymin = total_load_factor_growth - ci, ymax = total_load_factor_growth + ci)) +
  coord_flip()
```

Charlotte had significantly higher load factor growth than other cities

```{r}
pop_increases %>% group_by(origin_city) %>% 
  summarize(min_year = min(year), percent_pop_growth = sum(pop_increase, na.rm = TRUE) / avg_pop[year == min_year] * 100) %>% 
  arrange(desc(percent_pop_growth)) %>% 
  head(10) %>% 
  ggplot(mapping = aes(x = reorder(origin_city, percent_pop_growth), y = percent_pop_growth)) +
  geom_bar(stat = 'identity') +
  coord_flip()

```

## Look at plots of correlated vars

```{r}
#all points to show trend
yrly_com_flights_by_city %>% 
  ggplot(mapping = aes(x = total_pop_growth, y = total_passengers, label = origin_city)) +
  geom_point() 

#top 20 by poulation growth
yrly_com_flights_by_city %>% 
  arrange(desc(total_pop_growth)) %>% 
  head(20) %>% 
  ggplot(mapping = aes(x = total_pop_growth, y = total_passengers, label = origin_city)) +
  geom_point() +
  geom_label_repel()

yrly_com_flights_by_city %>% 
  arrange(desc(total_pop_growth)) %>% 
  head(20) %>% 
  ggplot(mapping = aes(x = total_pop_growth, y = total_seat_growth, label = origin_city)) +
  geom_point() +
  geom_label_repel()
```

```{r}
yrly_com_flights_by_city %>% 
  arrange(desc(percent_pop_growth)) %>% 
  head(20) %>% 
  ggplot(mapping = aes(x = percent_pop_growth, y = total_passenger_growth, label = origin_city)) +
  geom_point() +
  geom_label_repel()
```

```{r}
yrly_com_flights_by_city %>% 
  arrange(desc(percent_pop_growth)) %>% 
  head(20) %>% 
  ggplot(mapping = aes(x = percent_pop_growth, y = passenger_growth_percentage, label = origin_city)) +
  geom_point() +
  geom_label_repel()

yrly_com_flights_by_city %>% 
  arrange(desc(passenger_growth_percentage)) %>% 
  head(20) %>% 
  ggplot(mapping = aes(x = percent_pop_growth, y = passenger_growth_percentage, label = origin_city)) +
  geom_point() +
  geom_label_repel()
```

+ Remember, overall passenger volumes are down so Austin and Charlotte are outliers for growing

```{r}
yrly_com_flights_by_city %>% 
  arrange(desc(percent_pop_growth)) %>% 
  head(20) %>% 
  ggplot(mapping = aes(x = total_load_factor, y = percent_pop_growth, label = origin_city)) +
  geom_point() +
  geom_label_repel()

yrly_com_flights_by_city %>% 
  arrange(desc(total_load_factor)) %>% 
  head(20) %>% 
  ggplot(mapping = aes(x = total_load_factor, y = percent_pop_growth, label = origin_city)) +
  geom_point() +
  geom_label_repel()
```

```{r}
yrly_com_flights_by_city %>% 
  arrange(desc(total_pop_growth)) %>% 
  head(20) %>% 
  ggplot(mapping = aes(x = total_load_factor, y = total_pop_growth, label = origin_city)) +
  geom_point() +
  geom_label_repel()
```

```{r}
#load factor growths
yrly_com_flights_by_city %>% 
  arrange(desc(total_load_factor_growth)) %>% 
  head(20) %>% 
  ggplot(mapping = aes(x = total_load_factor_growth, y = total_pop_growth, label = origin_city)) +
  geom_point() +
  geom_label_repel()

yrly_com_flights_by_city %>% 
  arrange(desc(avg_load_factor_growth)) %>% 
  head(20) %>% 
  ggplot(mapping = aes(x = avg_load_factor_growth, y = total_pop_growth, label = origin_city)) +
  geom_point() +
  geom_label_repel()

#load factor and passenger growths
yrly_com_flights_by_city %>% 
  arrange(desc(avg_load_factor_growth)) %>% 
  head(20) %>% 
  ggplot(mapping = aes(x = avg_load_factor_growth, y = total_passenger_growth, label = origin_city)) +
  geom_point() +
  geom_label_repel()
```

```{r}
# all cities with positve passenger growth and positive load factor growth
yrly_com_flights_by_city %>% 
  filter(total_passenger_growth > 100000, avg_load_factor_growth > 0) %>% 
  arrange(desc(total_passenger_growth)) %>% 
  ggplot(mapping = aes(x = avg_load_factor_growth, y = total_passenger_growth, label = origin_city)) + 
  geom_point() +
  geom_label_repel()
  
```


# Summary of findings

+ Charlotte, NC has largest load factor growth, as well as largest passenger growth (as percentage)
    + Major positive outlier and must be candidate for further expansion
+ Dallas, TX has above average load factor and very large population growth
    + Implies room to expoand there as well
+ Other cities with population growth, passenger growth, and high load factor:
- Seattle, WA
- Austin, TX
- Phoenix, AZ


# Plots

## Start with general industry summary charts

```{r}
#save com flights df for using in other files
write_csv(com_flights_df, path = 'com_flights_df.csv')

pop_trend_data <- com_flights_df %>% 
  group_by(year) %>% 
  summarise(tot_passenger_vol = sum(passengers))

pop_trend_plot <- ggplot(pop_trend_data, mapping = aes(x = year, y = tot_passenger_vol/1e6)) +
  geom_line(color = 'grey') +
  geom_line(filter(pop_trend_data, year > 2006), 
            mapping = aes(x = year, y = tot_passenger_vol/1e6),
            size = 1.5, 
            color = '#750000')

pop_trend_plot <- pop_trend_plot + 
  geom_point(data = filter(pop_trend_data, year == 2009), 
             mapping = aes(x = year, y = tot_passenger_vol/1e6), 
             fill = '#344182', shape = 25, size = 3) +
  geom_text(aes(x = 2010, y = tot_passenger_vol[year == 2009]/1e6 - 10), label = '526.3M', color = '#344182' )

(pop_trend_data$tot_passenger_vol[20] - pop_trend_data$tot_passenger_vol[18]) / 
  pop_trend_data$tot_passenger_vol[18] * 100

pop_trend_plot + theme_classic() +
  theme(axis.line.y = element_blank(),
        axis.line.x = element_blank(),
        axis.title.x  = element_blank(),
        plot.title = element_text(color = "#808080",
                                  size = 16, face = "bold"),
        plot.subtitle = element_text(color = "#808080", 
                                     size = 12, face = "plain"),
        axis.text.x = element_text(color = "#808080", size = 11),
        axis.text.y = element_text(color = "#808080", size = 11),
        axis.title = element_text(color = "#808080", size = 11, face = "plain"),
        legend.title = element_text(color = "#808080", size = 12, face = "plain"),
        legend.text = element_text(color = "#808080", size = 12, face = "plain"),
        plot.caption = element_text(color = "#808080", size = 10, face = "plain", margin = margin(t=10))
        ) +
  scale_x_continuous(breaks = seq(2005,2010, by = 2)) +
  labs(title = 'Total passenger volume is falling',
       subtitle = expression(paste('Down ',
                                   italic(bold('9.1%')),
                                   ' from 2007 peak')),
       y = 'Millions of Passengers')
```

```{r}
filled_trend_data <- com_flights_df %>% 
  group_by(year) %>% 
  summarise(avg_load_factor = mean(load_factor))

filled_trend_plot <- ggplot(filled_trend_data, 
                            mapping = aes(x = year,
                                          y = avg_load_factor/100)) +
  geom_line(color = 'grey') +
  geom_line(filter(filled_trend_data, year > 2006), 
            mapping = aes(x = year, y = avg_load_factor/100),
            size = 1.5, 
            color = '#344182')

filled_trend_plot <- filled_trend_plot + 
  geom_point(data = filter(filled_trend_data, year == 2009), 
             mapping = aes(x = year, y = avg_load_factor/100), 
             fill = '#344182', shape = 25, size = 3) +
  geom_text(aes(x = 2010, y = avg_load_factor[year == 2009]/100 - 0.01), label = '71%', color = '#344182' )


filled_trend_plot + theme_classic() +
  theme(axis.line.y = element_blank(),
        axis.line.x = element_blank(),
        axis.title  = element_blank(),
        plot.title = element_text(color = "#808080",
                                  size = 16, face = "bold"),
        plot.subtitle = element_text(color = "#808080", 
                                     size = 12, face = "plain"),
        axis.text.x = element_text(color = "#808080", size = 11),
        axis.text.y = element_text(color = "#808080", size = 11),
        legend.title = element_text(color = "#808080", size = 12, face = "plain"),
        legend.text = element_text(color = "#808080", size = 12, face = "plain"),
        plot.caption = element_text(color = "#808080", size = 10, face = "plain", margin = margin(t=10))
        ) +
  scale_x_continuous(breaks = seq(2005,2010, by = 2)) +
  scale_y_continuous(labels = percent_format()) +
  labs(title = 'However, percent of seats filled has gone up',
       subtitle = 'Up 1 percentage point from 2007 total passenger peak')
```

## Recommendation Charts
```{r}
yrly_com_flights_by_city %>% 
  ggplot(mapping = aes(x = total_pop_growth/1000000, y = total_passengers/1000000)) +
  geom_point(color = 'grey') + 
  geom_smooth(color = '#344182', method = 'lm', se = FALSE) +
  theme_classic() +
  theme(axis.line.y = element_blank(),
        axis.line.x = element_blank(),
        axis.title  = element_text(color = "#808080", size = 11, face = "bold"),
        plot.title = element_text(color = "#808080",
                                  size = 16, face = "bold"),
        plot.subtitle = element_text(color = "#808080", 
                                     size = 12, face = "plain"),
        axis.text.x = element_text(color = "#808080", size = 11),
        axis.text.y = element_text(color = "#808080", size = 11),
        legend.title = element_text(color = "#808080", size = 12, face = "plain"),
        legend.text = element_text(color = "#808080", size = 12, face = "plain"),
        plot.caption = element_text(color = "#808080", size = 10, face = "plain", margin = margin(t=10))
        ) +
  xlab('Population growth') +
  ylab('Total passengers') +
  labs(title = 'Population growth correlates strongly with passenger growth',
       subtitle = 'Numbers for previous five years (in millions)\nDallas, TX fastest growing city (by far)') +
  geom_label_repel(filter(yrly_com_flights_by_city, origin_city == 'Dallas, TX'), 
                   mapping = aes(x = total_pop_growth/1000000, y = total_passengers/1000000, label = origin_city),
                   segment.color = '#344182', min.segment.length = 0, nudge_y = -15.5, label.size = NA, color = '#344182')
```

```{r}
# all cities with positve passenger growth and positive load factor growth
yrly_com_flights_by_city %>% 
  filter(total_passenger_growth > 0, avg_load_factor_growth > 0) %>% 
  arrange(desc(total_passenger_growth)) %>% 
  ggplot(mapping = aes(x = avg_load_factor_growth/100, y = total_passenger_growth/1e6, label = origin_city)) + 
  geom_point(color = 'grey') + 
  theme_classic() +
  theme(axis.line.y = element_blank(),
        axis.line.x = element_blank(),
        axis.title  = element_text(color = "#808080", size = 11, face = "bold"),
        plot.title = element_text(color = "#808080",
                                  size = 16, face = "bold"),
        plot.subtitle = element_text(color = "#808080", 
                                     size = 12, face = "plain"),
        axis.text.x = element_text(color = "#808080", size = 11),
        axis.text.y = element_text(color = "#808080", size = 11),
        legend.title = element_text(color = "#808080", size = 12, face = "plain"),
        legend.text = element_text(color = "#808080", size = 12, face = "plain"),
        plot.caption = element_text(color = "#808080", size = 10, face = "plain", margin = margin(t=10))
        ) +
  scale_x_continuous(labels = percent_format()) +
  xlab('Annualized load factor growth') +
  ylab('Passenger growth (in millions)') +
  labs(title = 'Charlotte is a major positive growth outlier',
       subtitle = 'Both load factor growth and passenger growth are strong\n(only positve passenger growth cities shown) ') +
  geom_label_repel(filter(yrly_com_flights_by_city, origin_city %in% c('Charlotte, NC', 'Seattle, WA', 'Miami, FL')), 
                   mapping = aes(x = avg_load_factor_growth/100, y = total_passenger_growth/1e6, label = origin_city),
                   segment.color = '#344182', min.segment.length = 0, nudge_y = -.15, label.size = NA, color = '#344182')

```

```{r}
yrly_com_flights_by_city %>% 
  filter(percent_pop_growth > 0) %>% 
  filter(total_load_factor > 0.65) %>% 
  ggplot(mapping = aes(x = total_load_factor, y = percent_pop_growth/100)) +
  geom_point(color = 'grey') +
  theme_classic() +
  theme(axis.line.y = element_blank(),
        axis.line.x = element_blank(),
        axis.title  = element_text(color = "#808080", size = 11, face = "bold"),
        plot.title = element_text(color = "#808080",
                                  size = 16, face = "bold"),
        plot.subtitle = element_text(color = "#808080", 
                                     size = 12, face = "plain"),
        axis.text.x = element_text(color = "#808080", size = 11),
        axis.text.y = element_text(color = "#808080", size = 11),
        legend.title = element_text(color = "#808080", size = 12, face = "plain"),
        legend.text = element_text(color = "#808080", size = 12, face = "plain"),
        plot.caption = element_text(color = "#808080", size = 10, face = "plain", margin = margin(t=10))
        ) +
  scale_x_continuous(labels = percent_format()) +
  scale_y_continuous(labels = percent_format()) +
  xlab('Load factor') +
  ylab('Population growth') +
  labs(title = 'Seattle and Miami are also prime candidates for expansion',
       subtitle = 'High load factor, above average population growth') +
  geom_label_repel(filter(yrly_com_flights_by_city, origin_city %in% c('Charlotte, NC', 'Seattle, WA', 'Dallas, TX', 'Miami, FL')), 
                   mapping = aes(x = total_load_factor, y = percent_pop_growth/100, label = origin_city),
                   segment.color = '#344182', min.segment.length = 0, nudge_y = 0.05, nudge_x = 0.05, label.size = NA, color = '#344182')
```


