---
title: "Flight Data Analysis"
author: "Eric Hestekin"
date: "2/27/2020"
output: 
  html_document:
    number_sections: yes
    toc: yes
    toc_depth: 3
    toc_float: yes
---

Repo:  https://github.com/ehestekin/MSBA_5210_DataVis

``` {r echo = FALSE}
# Course: 5210 Communicating Data
# Purpose: General Mills Cereal Sales Analysis
# Date: 2/13/20
# Author: Eric Hestekin, Schyuler Lujan
# Repo:  https://github.com/ehestekin/MSBA_5210_DataVis
```

``` {r echo = FALSE, include = FALSE}
# Clear environment of variables and functions
rm(list = ls(all = TRUE)) 
# Clear environmet of packages
if(is.null(sessionInfo()$otherPkgs) == FALSE)lapply(paste("package:", names(sessionInfo()$otherPkgs), sep=""), detach, character.only = TRUE, unload = TRUE)
```

``` {r setup, include = FALSE, warning = FALSE}
# Load Library
library(tidyverse)
library(patchwork)
library(scales)
```


# Intro

Data source :  [US domestic flights from 1990 to 2009, US Census Bureau](http://academictorrents.com/details/a2ccf94bbb4af222bf8e69dad60a68a29f310d9a)

```{r, warning = FALSE}
#Load Data TSV
#no column headers so set array of names
col_labels = c('origin_code',
               'dest_code',
               'origin_city',
               'dest_city',
               'passengers',
               'seats',
               'num_flights',
               'distance',
               'year_month',
               'origin_pop',
               'dest_pop')

flights_df <- read_tsv('flight_edges.tsv',col_names = col_labels)
```

Interested in looking at 
+ passenger volume
+ percent capacity
+ regional trends
+ total flights

# Basic EDA

## First look at data

```{r}
head(flights_df)

summary(flights_df)
```

## Tidy

```{r}
#separate month_year
#first 4 chars are year, convert from string to num
flights_df <- flights_df %>% 
  separate(year_month, into = c('year','month'), sep = 4, convert = TRUE) 

#add percentage of seats filled for each flight
#and remove rows where seats are zero (no flight data)
# and also remove flights where seat/flight < 4 (non commercial flight)
com_flights_df <- flights_df %>%
  mutate(seat_fill_perc = (passengers/seats) * 100) %>% 
  filter(seats != 0) %>% 
  filter(seats/num_flights > 4)

#also throw out seat filled percentages > 100
#this shouldn't be possible so don't trust data point
com_flights_df <- filter(com_flights_df, seat_fill_perc <= 100)

#get summary now with tidy commercial flights data set
summary(com_flights_df)
```

*Notes*
+ Would be useful to look at averages for passengers and seats on a single flight
+ Interesting distribution of distances
 + + Looks like many short routes (median << mean)
+ Population skewed towards very large cities (due to more flight routes?)
+ Fairly symmetrical seat filled percentage

## Look at year to year trends

```{r}
com_flights_df %>% 
  group_by(year) %>% 
  summarise(avg_num_flights = mean(num_flights)) %>% 
  ggplot(mapping = aes(x = year, y = avg_num_flights)) +
  geom_line() +
  geom_point()

com_flights_df %>% 
  group_by(year) %>% 
  summarise(avg_seat_fill_perc = mean(seat_fill_perc)) %>% 
  ggplot(mapping = aes(x = year, y = avg_seat_fill_perc)) +
  geom_line() +
  geom_point()

com_flights_df %>% 
  group_by(year) %>% 
  summarise(avg_passenger_vol = mean(passengers)) %>% 
  ggplot(mapping = aes(x = year, y = avg_passenger_vol)) +
  geom_line() +
  geom_point()

com_flights_df %>% 
  group_by(year) %>% 
  summarise(tot_passenger_vol = sum(passengers)) %>% 
  ggplot(mapping = aes(x = year, y = tot_passenger_vol)) +
  geom_line() +
  geom_point()
```

+ Drop in passenger volumes from 2007 - 2009
+ Steady increase in percentage of seats filled


## Look at individual routes

```{r}
route_data <- com_flights_df %>% 
  group_by(origin_city, dest_city, year) %>% 
  summarise(total_flights = sum(num_flights),
            total_passengers = sum(passengers),
            total_seats = sum(seats),
            avg_seat_fill_perc = mean(total_passengers / total_seats) * 100)

#lets look at 5 most popular routes from last year and see how it has trended over the previous 10 years
#only look at routes with more than a flight every other day
route_data_2009 <- route_data %>% filter(year == 2009, total_flights > 365/2)

#look at 10 most passenger routes
head(route_data_2009 %>% arrange(desc(total_passengers)), 10)
#and most and least filled
head(route_data_2009 %>% arrange(desc(avg_seat_fill_perc)), 10)
head(route_data_2009 %>% arrange(avg_seat_fill_perc), 10)


```

```{r}
best_filled_routes_2009 <- route_data_2009 %>%
  arrange(desc(avg_seat_fill_perc)) %>% 
  select(origin_city,dest_city)

#had a hard time subsetting first x rows with pipe and filter/slice
#easy indexing here
top10_filled_routes_2009 <- best_filled_routes_2009[1:10,]

#similar with most popular
most_pop_routes_2009 <- route_data_2009 %>% 
  arrange(desc(total_passengers)) %>% 
  select(origin_city,dest_city)

top10_pop_routes_2009 <- most_pop_routes_2009[1:10,]
```

Plot the above popular and most filled routes over previous five years

```{r}
com_flights_df %>% 
  filter(origin_city %in% top10_filled_routes_2009$origin_city &
           dest_city %in% top10_filled_routes_2009$dest_city) %>% 
  group_by(origin_city,dest_city,year) %>% 
  summarize(avg_seat_fill_perc = 
              mean(sum(passengers) / sum(seats)) * 100) %>% 
  ggplot(mapping = aes(x = year, y = avg_seat_fill_perc, color = origin_city)) +
  geom_point()

#plot is too busy and having trouble filtering routes liked i'd like
#manually choose origin/dest pairs
com_flights_df %>% 
  filter(origin_city == 'Nashville, TN' &
           dest_city == 'Seattle, WA') %>% 
  group_by(origin_city,dest_city,year) %>% 
  summarize(avg_seat_fill_perc = 
              mean(sum(passengers) / sum(seats)) * 100) %>% 
  ggplot(mapping = aes(x = year, y = avg_seat_fill_perc, color = origin_city)) +
  geom_point()
```


## Look at city/population relations

```{r}
com_flights_by_city <- com_flights_df %>% 
  group_by(origin_code, year) %>% 
  summarize(tot_passenger_departs = sum(passengers),
            origin_pop = mean(origin_pop), #still want origin pop data.  mean should just return the pop
            avg_seat_fill = sum(passengers)/sum(seats) * 100,
            pass_pop_ratio = tot_passenger_departs / origin_pop) %>% 
  left_join(distinct(select(com_flights_df, origin_code, origin_city)), by = 'origin_code')

#limit to cities with significant departures (>100 passengers daily)
com_flights_by_city <- com_flights_by_city %>% filter(tot_passenger_departs > 36500)

com_flights_by_city_2009 <- com_flights_by_city %>% filter(year == 2009)

summary(com_flights_by_city)
```

# Plots

```{r}
pop_trend_data <- com_flights_df %>% 
  group_by(year) %>% 
  summarise(tot_passenger_vol = sum(passengers))

pop_trend_plot <- ggplot(pop_trend_data, mapping = aes(x = year, y = tot_passenger_vol/1e6)) +
  geom_line(color = 'grey') +
  geom_line(filter(pop_trend_data, year > 2006), 
            mapping = aes(x = year, y = tot_passenger_vol/1e6),
            size = 1.5, 
            color = '#750000')

pop_trend_plot <- pop_trend_plot + 
  geom_point(data = filter(pop_trend_data, year == 2009), 
             mapping = aes(x = year, y = tot_passenger_vol/1e6), 
             fill = '#344182', shape = 25, size = 3) +
  geom_text(aes(x = 2010, y = tot_passenger_vol[year == 2009]/1e6 - 10), label = '526.3M', color = '#344182' )

(pop_trend_data$tot_passenger_vol[20] - pop_trend_data$tot_passenger_vol[18]) / 
  pop_trend_data$tot_passenger_vol[18] * 100

pop_trend_plot + theme_classic() +
  theme(axis.line.y = element_blank(),
        axis.line.x = element_blank(),
        axis.title.x  = element_blank(),
        plot.title = element_text(color = "#808080",
                                  size = 16, face = "bold"),
        plot.subtitle = element_text(color = "#808080", 
                                     size = 12, face = "plain"),
        axis.text.x = element_text(color = "#808080", size = 11),
        axis.text.y = element_text(color = "#808080", size = 11),
        axis.title = element_text(color = "#808080", size = 11, face = "plain"),
        legend.title = element_text(color = "#808080", size = 12, face = "plain"),
        legend.text = element_text(color = "#808080", size = 12, face = "plain"),
        plot.caption = element_text(color = "#808080", size = 10, face = "plain", margin = margin(t=10))
        ) +
  scale_x_continuous(breaks = seq(1990,2010, by = 6)) +
  labs(title = 'Total passenger volume is falling',
       subtitle = expression(paste('Down ',
                                   italic(bold('9.1%')),
                                   ' from 2007 peak')),
       y = 'Millions of Passengers')


```


```{r}
filled_trend_data <- com_flights_df %>% 
  group_by(year) %>% 
  summarise(avg_seat_fill_perc = mean(seat_fill_perc))

filled_trend_plot <- ggplot(filled_trend_data, 
                            mapping = aes(x = year,
                                          y = avg_seat_fill_perc/100)) +
  geom_line(color = 'grey') +
  geom_line(filter(filled_trend_data, year > 2006), 
            mapping = aes(x = year, y = avg_seat_fill_perc/100),
            size = 1.5, 
            color = '#344182')

filled_trend_plot <- filled_trend_plot + 
  geom_point(data = filter(filled_trend_data, year == 2009), 
             mapping = aes(x = year, y = avg_seat_fill_perc/100), 
             fill = '#344182', shape = 25, size = 3) +
  geom_text(aes(x = 2010, y = avg_seat_fill_perc[year == 2009]/100 - 0.01), label = '71%', color = '#344182' )


filled_trend_plot + theme_classic() +
  theme(axis.line.y = element_blank(),
        axis.line.x = element_blank(),
        axis.title  = element_blank(),
        plot.title = element_text(color = "#808080",
                                  size = 16, face = "bold"),
        plot.subtitle = element_text(color = "#808080", 
                                     size = 12, face = "plain"),
        axis.text.x = element_text(color = "#808080", size = 11),
        axis.text.y = element_text(color = "#808080", size = 11),
        legend.title = element_text(color = "#808080", size = 12, face = "plain"),
        legend.text = element_text(color = "#808080", size = 12, face = "plain"),
        plot.caption = element_text(color = "#808080", size = 10, face = "plain", margin = margin(t=10))
        ) +
  scale_x_continuous(breaks = seq(1990,2010, by = 6)) +
  scale_y_continuous(labels = percent_format()) +
  labs(title = 'However, percent of seats filled has gone up',
       subtitle = 'Up 1 percentage point from 2007 total passenger peak')
```

```{r}
com_flights_by_city_2009 %>% 
  arrange(desc(tot_passenger_departs)) %>% 
  filter(tot_passenger_departs > 16500000) %>% 
  select(origin_code, tot_passenger_departs, origin_city) %>%
  ggplot(mapping = aes(x = origin_code, y = tot_passenger_departs)) +
  geom_bar(stat = 'identity') +
  coord_flip()
```

```{r}
com_flights_by_city_2009 %>% 
  arrange(desc(tot_passenger_departs)) %>% 
  filter(tot_passenger_departs > 13350000) %>% 
  select(origin_code, origin_city, tot_passenger_departs) %>%
  mutate(tot_passenger_departs = tot_passenger_departs / 1e6) %>% 
  knitr::kable(col.names = c('Airport Code',
                             'City',
                             'Passengers (Millions)'),
               align = 'c',
               digits = 2) %>% 
  kableExtra::row_spec(1)
```

```{r}
com_flights_by_city_2009 %>% 
  arrange(desc(avg_seat_fill)) %>% 
  filter(tot_passenger_departs > 1e6,
         avg_seat_fill > 79.475) %>% 
  select(origin_code, origin_city, avg_seat_fill) %>%
  knitr::kable(col.names = c('Airport Code',
                             'City',
                             'Percent Seats Filled'),
               align = 'c',
               digits = 2) %>% 
  kableExtra::row_spec(8, bold = T, 
                       color = '#808080', background = '#344182')
```

```{r}
route_data_2009 %>% 
  arrange(desc(avg_seat_fill_perc)) %>% 
  filter(avg_seat_fill_perc > 94.4) %>% 
  select(origin_city, dest_city, avg_seat_fill_perc) %>%
  ungroup() %>% 
  add_row(origin_city = 'All Routes',
                     dest_city = 'Median:',
                     avg_seat_fill_perc = 75.284) %>% 
  knitr::kable(col.names = c('Origin',
                             'Destination',
                             'Percent Seats Filled'),
               align = rep('c', 3),
               digits = 2) %>% 
  kableExtra::row_spec(c(1,3), bold = T, 
                       color = '#808080', background = '#344182') %>% 
  kableExtra::kable_styling(full_width = T) %>% 
  kableExtra::row_spec(4, bold = T)

# summary(route_data_2009)
```

```{r}
route_data_2009 %>% 
  arrange(desc(total_passengers)) %>% 
  filter(total_passengers %in% 
           #manually select rows for now had a hard time filtering properly
           c(1501883, 1494141, 1399554, 1380928, 1352360)) %>% 
  mutate(route = paste(origin_city, ' - ', dest_city)) %>% 
  ggplot(mapping = aes(x = route, y = avg_seat_fill_perc/100)) +
  geom_bar(stat = 'identity', 
           fill = c('grey','#750000','grey','grey','grey')) +
  geom_text(aes(x = route, y = avg_seat_fill_perc/100,
                label = scales::percent(avg_seat_fill_perc/100,
            accuracy = 1),
            fontface = c('plain','plain','plain','bold','plain')), 
            position = position_dodge(width = 0.9), 
            hjust = 1) +
  coord_flip() + theme_classic() +
  theme(axis.line.y = element_blank(),
        axis.line.x = element_blank(),
        axis.title  = element_blank(),
        plot.title = element_text(color = "#808080",
                                  size = 16, face = "bold", hjust = -0.25),
        plot.subtitle = element_text(color = "#808080", 
                                     size = 12, face = "plain"),
        axis.text.x = element_blank(),
        axis.ticks = element_blank(),
        axis.text.y = element_text(color = "#808080", size = 11,
                                   face = c('plain','bold','plain','plain','plain')),
        legend.title = element_text(color = "#808080", size = 12, face = "plain"),
        legend.text = element_text(color = "#808080", size = 12, face = "plain"),
        plot.caption = element_text(color = "#808080", size = 10, face = "plain", margin = margin(t=10))
        ) +
  scale_y_continuous(labels = percent_format(accuracy = 1)) +
  labs(title = 'Houston to Dallas flights are too empty',
       subtitle = 'Top 5 routes by passenger volume' )
  
```

